Fundamental Problems with the Evidence Base of Adolescent Depression Treatments

### 2024-04-12 Manuscript_for_rendering_in_word.qmd
Test

###########Libraries needed
library(tidyverse)
library(truncnorm)
library(metafor)
library(meta)
library(kableExtra)
library(flextable)
library(magrittr)
library(officer)
library(RefManageR)

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
library(tidyverse)
# Transforming data ------------------------------------------------------
# import master dataset

@@

# Get Ns, run metareg, prepare graphing ------------------------------------------------------
library(tidyverse)
library(metafor)

@@

flextable(df_overall_baseline_results) %>%
  set_caption("Sample characteristics at baseline across medication and psychotherapy studies") %>%
  bold(part = "header") %>%  # Bold the header row
  bold(i = c(1, 4, 7), j = NULL, bold = TRUE, part = "body") %>%  # Bold specific rows
  border_outer(border = small_border) %>%  # Add outer borders
  hline(i = c(1, 2, 4, 5, 7, 8), border = small_border) %>%  # Add light grey horizontal lines all  rows
  hline(i = c(3, 6), border = big_border) %>%  # Add darker horizontal lines after specific rows %>% 
  hline(part = "header", border = big_border) %>% 
  vline(j = 1, border = big_border) %>% 
  set_table_properties(width = .75) %>%  # Set table width
  bg(bg = "#F2F2F2", part = "header") %>%  # Add light grey background to header row
  autofit() %>%  # Autofit column widths
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri 
  padding(i = c(2, 3, 5, 6, 8, 9), j = 1, padding.left = 10) %>% 
  align(j = 2:ncol(df_overall_baseline_results), align = "center", part = "all")

@@

@@

### Number of trial sites

@@

df_baseline_demographics %>%
  group_by(psy_or_med, site_binary) %>%
  summarize(num_trials = n())

@@

flextable(df_sites_summary_stats) %>%
  set_caption ("Number of sites across medication and psychotherapy studies", style = "Table Caption") %>%
  bold(part = "header") %>%  
  border_outer(border = small_border) %>%  
  hline(part = "all", border = small_border) %>% 
  hline(part = "header", border = big_border) %>% 
  vline(j = 1, border = big_border) %>% 
  set_table_properties(width = .75) %>% 
  bg(bg = "#F2F2F2", part = "header") %>%  
  autofit() %>% 
  font(fontname = "Calibri", part = "all") %>%   
  padding(i = c(1, 2), j = 1, padding.left = 10) %>% 
  align(j = 2:ncol(df_sites_summary_stats), align = "center", part = "all")

@@

There was a significant difference between the number of sites in medication versus psychotherapy trials, as can be seen in Tables 2 and 3. Average number of trial sites was significantly higher across medication studies (*M* = `r df_sites_summary_stats$Mean[df_sites_summary_stats$Modality == "Medication"]`, *SD* =`r df_sites_summary_stats$SD[df_sites_summary_stats$Modality == "Medication"]`) compared to psychotherapy studies (*M* =`r df_sites_summary_stats$Mean[df_sites_summary_stats$Modality == "Psychotherapy"]`, *SD* =`r df_sites_summary_stats$SD[df_sites_summary_stats$Modality == "Psychotherapy"]`)(*t* (`r round (sites_t_test$parameter, 2)`) = `r round(sites_t_test$statistic,2)`, *p* =`r round(sites_t_test$p.value, 3`). Of those studies for which we had data on number of sites, 26 of 28 (93%) medication trials were multisite, compared to 24 of 45 (54%) psychotherapy studies.

## Mean efficacy of psychotherapy versus medication control conditions

@@

@@

Tables 6a and 6b show the baeline-mean adjusted SMDs for the CDRS at outcome when the baseline measurement has been taken into account. <!--# Here Intercept refers to teh psychotherapy control. Sorry about the labelling of the rows, will improve. -->

>>>>

###########Libraries needed
library(tidyverse)
library(truncnorm)
library(metafor)
library(meta)
library(kableExtra)
library(flextable)
library(magrittr)
library(officer)
library(RefManageR)
library(gt)
library(officer)
library(officedown)

@@

# Transforming data ------------------------------------------------------
# import master dataset

@@

# Get Ns, run metareg, prepare graphing ------------------------------------------------------

@@

# knitr:: kable(df_overall_baseline_results, caption="Sample characteristics at baseline across medication and psychotherapy studies")
# 
# gt(df_overall_baseline_results) %>% 
#   tab_header(
#     title = md("Test this"),
#     subtitle = md("And this"))

@@

```{r, echo=FALSE}

#| label: tbl-flextable
#| tbl-cap: Caption **strong**
flextable:: flextable(df_overall_baseline_results) %>%
  set_caption("Sample characteristics at baseline across medication and psychotherapy studies") %>%
  bold(part = "header") %>%  # Bold the header row
  bold(i = c(1, 4, 7), j = NULL, bold = TRUE, part = "body") %>%  # Bold specific rows
  border_outer(border = small_border) %>%  # Add outer borders
  hline(i = c(1, 2, 4, 5, 7, 8), border = small_border) %>%  # Add light grey horizontal lines all  rows
  hline(i = c(3, 6), border = big_border) %>%  # Add darker horizontal lines after specific rows %>% 
  hline(part = "header", border = big_border) %>% 
  vline(j = 1, border = big_border) %>% 
  set_table_properties(width = .75) %>%  # Set table width
  bg(bg = "#F2F2F2", part = "header") %>%  # Add light grey background to header row
  autofit() %>%  # Autofit column widths
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri 
  padding(i = c(2, 3, 5, 6, 8, 9), j = 1, padding.left = 10) %>% 
  align(j = 2:ncol(df_overall_baseline_results), align = "center", part = "all")

@@

## Number of trial sites

@@

# df_baseline_demographics %>%
#   group_by(psy_or_med, site_binary) %>%
#   summarize(num_trials = n())

@@

flextable(df_sites_summary_stats) %>%
  set_caption ("Number of sites across medication and psychotherapy studies") %>%
  bold(part = "header") %>%
  border_outer(border = small_border) %>%
  hline(part = "all", border = small_border) %>%
  hline(part = "header", border = big_border) %>%
  vline(j = 1, border = big_border) %>%
  set_table_properties(width = .75) %>%
  bg(bg = "#F2F2F2", part = "header") %>%
  autofit() %>%
  font(fontname = "Calibri", part = "all") %>%
  padding(i = c(1, 2), j = 1, padding.left = 10) %>%
  align(j = 2:ncol(df_sites_summary_stats), align = "center", part = "all")

@@

There was a significant difference between the number of sites in medication versus psychotherapy trials, as can be seen in Tables 2 and 3. Average number of trial sites was significantly higher across medication studies (*M* = `r df_sites_summary_stats$Mean[df_sites_summary_stats$Modality == "Medication"]`, *SD* =`r df_sites_summary_stats$SD[df_sites_summary_stats$Modality == "Medication"]`) compared to psychotherapy studies (*M* =`r df_sites_summary_stats$Mean[df_sites_summary_stats$Modality == "Psychotherapy"]`, *SD* =`r df_sites_summary_stats$SD[df_sites_summary_stats$Modality == "Psychotherapy"]`)(*t* (`r round (sites_t_test$parameter, 2)`) = `r round(sites_t_test$statistic,2)`, *p* =`r sites_t_test$p.value`). Of those studies for which we had data on number of sites, 26 of 28 (93%) medication trials were multisite, compared to 24 of 45 (54%) psychotherapy studies.

## Mean efficacy of control conditions in psychotherapy versus medication studies

@@

```{r, eval=FALSE} 
#### Let's start from the beginning and regress cohens on the baseline dataframe.

# create a new list with dfs for the sims
df_long_for_metan_adj_cohens_d <- df_long_for_metan
# rename the cohens_d value to make space for the new cohen's d
df_long_for_metan_adj_cohens_d <-
  df_long_for_metan_adj_cohens_d %>%
rename(unadjusted_cohens_d = cohens_d,
)
# estimate the cohen's d from the residuals
model_adjust_cohens_d <-  lm(df_long_for_metan_adj_cohens_d$unadjusted_cohens_d ~ df_long_for_metan_adj_cohens_d$perc_baseline_mean, na.action=na.exclude)

# there is a great effect of the baseline means, as shown here
results_adjusted_cohens_d <- broom:: tidy(model_adjust_cohens_d )

# extract the residuals which are the adjusted cohen's d, to get the simulated standard errors
df_long_for_metan_adj_cohens_d$cohens_d <- residuals(model_adjust_cohens_d )

# apply function to simulate standard errors
list_df_simulated_adj_cohens_d <- simulate_dataframes_for_st_errors(df = df_long_for_metan_adj_cohens_d,
                                   num_repetitions = 1000,
                                   seed  = 1974,
                                   n = length(unique(df_long_for_metan_adj_cohens_d$new_study_id)),
                                   a = 0.45,
                                   b = .9,
                                   mean = 0.65,
                                   sd = 0.2)

# add the four level variable and re-level
for(i in 1: length(list_df_simulated_adj_cohens_d)){

  list_df_simulated_adj_cohens_d[[i]]$arm_effect_size <- factor(list_df_simulated_adj_cohens_d[[i]]$arm_effect_size)
  list_df_simulated_adj_cohens_d[[i]] <- list_df_simulated_adj_cohens_d[[i]] %>%
    mutate(four_level_var = case_when(psy_or_med == 0 & arm_effect_size == "cohens_d_active" ~ "medication_active",
                                      psy_or_med == 0 & arm_effect_size == "cohens_d_control" ~ "medication_control",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_active" ~ "psychotherapy_active",
                                      psy_or_med == 1 & arm_effect_size == "cohens_d_control" ~ "psychotherapy_control")
    )

  list_df_simulated_adj_cohens_d[[i]]$four_level_var <- factor(list_df_simulated_adj_cohens_d[[i]]$four_level_var) # turn to factor

  # relevel so that medication control becomes the reference category for the regression
  #  list_df_simulated[[i]]$four_level_var <- relevel(list_df_simulated[[1]]$four_level_var, ref = "medication_control")
  list_df_simulated_adj_cohens_d[[i]]$four_level_var <-   fct_relevel(list_df_simulated_adj_cohens_d[[i]]$four_level_var,
                                                         "medication_control",
                                                         "medication_active",
                                                         "psychotherapy_control",
                                                         "psychotherapy_active")
}

# Run metaregression model (assuming that the four level variable is present given that I create this dataframe from the list_)
# specify model
model_1 <- as.formula(~ four_level_var)

# run metareg function
# here you can use the simulated results directly
list_model_1_meta_reg_adj_cohens_d <- run_metaregression(list_df_simulated_adj_cohens_d, model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated_adj_cohens_d[[1]]$four_level_var)
aggregate_results_overall_adjusted_cohens_d <-  aggregate_model_results(list_model_1_meta_reg_adj_cohens_d, condition)
aggregate_results_overall_adjusted_cohens_d <- cbind(aggregate_results_overall_adjusted_cohens_d, condition)

# extract SMDs from the  list
list_dummy_var_means_adjusted_cohens_d <-  lapply(list_model_1_meta_reg_adj_cohens_d, extract_coefficients_func)

# calculate mean SMDs and ses
df_mean_coefs_from_sim_adjusted_cohens_d <- calculate_mean_coefs_ses(list_dummy_var_means_adjusted_cohens_d, condition)

df_mean_coefs_from_sim_adjusted_cohens_d <- data.frame(cbind(condition, df_mean_coefs_from_sim_adjusted_cohens_d))

results_adjusted_cohens_d

@@

Tables 6a and 6b show the baseline-mean adjusted SMDs for the CDRS at outcome when the baseline measurement has been taken into account. <!--# Here Intercept refers to teh psychotherapy control. Sorry about the labelling of the rows, will improve. -->
